\textbf{ Discovery of a disrupting open cluster far into the Milky Way halo:\\ a recent star formation event in the leading arm of the Magellanic stream?}

### 2018-11-13 ms.tex
caption and abs

We report the discovery of a young ($\tau \sim \clage$), low-mass ($M \sim 1200~\msun$), metal-poor ($\feh \sim \clfeh$) stellar association in the Milky Way halo: \clustername.
From isochrone fitting to deeper, follow-up \decam\ imaging of the cluster, we also measure its distance, $D \approx \cldist$.
At its present Galactocentric cylindrical position $(R, z) \sim (23, 15)~\textrm{kpc}$, the association is (on the sky) near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center, on the other side of the Milky Way disk relative to the LMC.
If we assume that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the leading arm of the Magellanic stream.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with Milky Way disk.
The estimated age of \clustername\ is consistent with the time of last passage of the leading arm gas through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream has low column density, the Milky Way disk at this large radius has low gas density, and the relative velocity of the leading arm gas and Milky Way gas is large. However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

@@

% Notebook: Figure-CMDs
\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{figures/gaia-cmd-pm.pdf}
\caption{Color-magnitude diagram, sky positions, and proper motions from \gaia\ \DR{2} for the region around $(\alpha, \delta) \sim (179, -29)^\circ$.
\textbf{Left panel}: \gaia\ color-magnitude diagram (CMD) for all sources with $G < 20$ and sky position within 3 degrees of $(\alpha, \delta) = (179.5, -28.8)^\circ$, extinction corrected following \citet{Danielski:2018}.
Red line shows a $100~\textrm{Myr}$, $\feh = -1.1$ \acronym{MIST} isochrone shifted to a distance of $30~\kpc$.
Shaded pixels (orange) show regions of high density and points (black markers) are plotted when there are fewer than 8 stars per pixel.
\textbf{Middle panel}: Sky positions for all sources in the shown sky region (grey 2D histogram).
Black markers show only sources in the blue selection polygon shown in the CMD (left panel).
\textbf{Right panel}: Proper motions for all sources in the shown sky region (grey 2D histogram).
Black markers again show only sources in the blue CMD selection region (left panel).
}
\label{fig:cmds}
\end{figure*}

>>>>

We report the discovery of a young ($\tau \sim \clage$), low-mass ($M \sim 1200~\msun$), metal-poor ($\feh \sim \clfeh$) stellar association in the Milky Way halo: \clustername.
From isochrone fitting to follow-up \decam\ imaging of the cluster, we also measure its distance, $D \approx \cldist$.
At its present Galactocentric position $(R, z) \sim (23, 15)~\textrm{kpc}$, the association is (on the sky) near the leading arm of the gas stream emanating from the Magellanic cloud system, but is located $\approx 60^\circ$ from the Large Magellanic Cloud (LMC) center on the other side of the Milky Way disk.
If we assume that the cluster is co-located with \hi gas in the stream, we directly measure the distance to the leading arm of the Magellanic stream.
At this location relative to the LMC, the measured distance is inconsistent with predictions from models of the LMC/SMC interaction and infall into the Milky Way that do not account for ram pressure and gas interaction with Milky Way disk.
The estimated age of \clustername\ is consistent with the time of last passage of the leading arm gas through the Galactic midplane, and we therefore speculate that this star-formation event was triggered by the last disk midplane passage, which occurred at a Galactocentric radius $R \approx 30~\kpc$.
However, most details of this idea remain a mystery: the Magellanic stream has low column density, the Milky Way disk at this large radius has low gas density, and the relative velocity of the leading arm gas and Milky Way gas is large. However it formed, the discovery of a young stellar cluster in the Milky Way halo presents an interesting opportunity for study.
\end{abstract}

@@

% Notebook: Figure-CMDs
\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{figures/gaia-cmd-pm.pdf}
\caption{Color-magnitude diagram, sky positions, and proper motions from \gaia\ \DR{2} for the region around $(\alpha, \delta) \sim (179, -29)^\circ$.
\textbf{Left panel}: \gaia\ color-magnitude diagram (CMD) for all sources with $G < 20$ and sky position within 3 degrees of $(\alpha, \delta) = (179.5, -28.8)^\circ$, extinction corrected following \citet{Danielski:2018}.
Red line shows a $100~\textrm{Myr}$, $\feh = -1.1$ \acronym{MIST} isochrone shifted to a distance of $30~\kpc$.
Shaded pixels (grey 2D histogram) show the density of all sources in the CMD for this sky region, and points (black markers) show only sources in the blue selection polygon shown.
\textbf{Middle panel}: Sky positions for all sources in the shown sky region (grey 2D histogram).
Black markers show only sources in the blue selection polygon shown in the CMD (left panel).
\textbf{Right panel}: Proper motions for all sources in the shown sky region (grey 2D histogram).
Black markers again show only sources in the blue CMD selection region (left panel).
}
\label{fig:cmds}
\end{figure*}
